---
title: "Projects"
description: "Selected academic and analytical projects."
author: null
output:
  distill::distill_article:
    toc: false
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```


## Selected Analytical Work

This page presents an example of academic and analytical project work completed as part of my Economics and Politics studies. The project demonstrates my ability to work with data, interpret evidence, create clear visualisations, and communicate findings in an accessible way.

```{=html}
<div class="card-grid"> <div class="info-card"> <h3>UK Constituency Analysis</h3> <p>An analysis of variation in constituency electorates across the United Kingdom, using official election data and R-based visualisation.</p> </div> <div class="info-card"> <h3>Text Analysis Project</h3> <p>A text-based analytical project exploring political or policy-related material through structured methods and interpretation.</p> </div> <div class="info-card"> <h3>Quantitative Politics Portfolio</h3> <p>Project-based work developed through my Economics and Politics studies, with a focus on data, evidence, and communication.</p> </div> </div>
```

## Featured Project: UK Constituency Electorate Analysis

This project examines whether constituency electorates vary across different parts of the United Kingdom.

In representative democracies, constituencies are expected to provide broadly comparable levels of representation. If some constituencies have much larger electorates than others, then MPs may represent very different numbers of voters. This raises important questions about electoral geography, representation, and institutional design.

The aim of this project was to demonstrate my ability to:

- Work with official election data
- Clean and organise data using R
- Produce clear and interpretable visualisations
- Explain quantitative results in accessible language
- Connect data analysis to wider political and institutional questions

## Data and Method

The analysis uses official UK Parliament election data. I cleaned the dataset, grouped results by constituency, and compared constituency electorates across England, Scotland, Wales, and Northern Ireland.

```{r constituency-analysis, fig.width=8, fig.height=5}
library(readr)
library(dplyr)
library(stringr)
library(ggplot2)

# Load UK election data directly from official source
uk_csv <- "https://electionresults.parliament.uk/general-elections/6/candidacies.csv"
uk_raw <- read_csv(uk_csv, show_col_types = FALSE)

# Clean column names
names(uk_raw) <- names(uk_raw) %>%
  tolower() %>%
  str_replace_all("\\s+", "_")

# Prepare constituency-level data
uk <- uk_raw %>%
  filter(!`election_is_by-election`) %>%
  select(
    constituency = constituency_name,
    country = country_name,
    valid_votes = candidate_vote_count,
    electorate = electorate
  ) %>%
  group_by(country, constituency) %>%
  summarise(
    valid_votes = sum(valid_votes, na.rm = TRUE),
    electorate = max(electorate, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(is.finite(electorate), electorate > 0)

# Visualise constituency electorate size by country
ggplot(uk, aes(x = country, y = electorate, fill = country)) +
  geom_boxplot(alpha = 0.7) +
  labs(
    title = "Constituency Electorate Size by Country",
    subtitle = "Boxplots show median, spread, and outliers across UK constituencies",
    x = "Country",
    y = "Electorate"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    legend.position = "none"
  )
```

## Interpretation

The boxplot shows that constituency electorate sizes vary across the United Kingdom.

England shows noticeable variation in electorate size, meaning that some constituencies have larger electorates than others. Scotland and Wales also show variation, while Northern Ireland appears to have a narrower spread compared with some other parts of the UK.

Overall, the analysis shows that constituencies can differ in the number of registered voters they contain. This is relevant because constituency size can inform wider debates about representation, electoral boundaries, and democratic institutions.

## Skills Demonstrated

This project demonstrates my ability to:

- Use R for data cleaning and analysis
- Work with real-world public data
- Create clear visualisations using `ggplot2`
- Summarise and interpret quantitative results
- Communicate findings in a clear and accessible way
- Connect data analysis to wider political and institutional questions

## Project Reflection

This analysis strengthened my understanding of how quantitative methods can be used to examine political institutions. It also helped me practise turning raw data into a clear visual output and explaining the results in a way that is understandable to both technical and non-technical audiences.