---
title: "Text Analysis of UK Political Manifestos"
author: "Muhammad Ahmad Hassan"
date: "2026-01-27"
output:
  distill::distill_article:
    toc: true
    toc_depth: 2
description: "A portfolio showcasing data analysis and visualisation work."
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(quanteda)
library(readtext)
library(ggplot2)
library(tibble)

```

## Introduction

This section applies text analysis techniques to examine patterns in textual data. This example uses a real collection of UK political manifestos to demonstrate key text analysis methods such as tokenization and word frequency analysis.


## Data Preparation

This section loads and prepares the text data. Each manifesto is read, combined into a single text per document, and converted into a corpus object for analysis.

```{r prepare_texts}
manifesto_dir <- if (dir.exists("UK_Manifestos")) {
  "UK_Manifestos"
} else {
  "../UK_Manifestos"
}

files <- list.files(manifesto_dir, pattern = "\\.txt$", full.names = TRUE)

manifestos <- lapply(files, function(f) {
  paste(readLines(f, warn = FALSE, encoding = "UTF-8"), collapse = " ")
})

texts <- as.character(unlist(manifestos))
doc_names <- basename(files)

manifestos_corpus <- quanteda::corpus(texts, docnames = doc_names)

## Token Cleaning

tokens_clean <- quanteda::tokens(
  manifestos_corpus,
  remove_punct = TRUE,
  remove_numbers = TRUE
) |>
  quanteda::tokens_remove(stopwords("en"))

dfm_mat <- quanteda::dfm(tokens_clean)
```


```{r top_words}
## Top Words in UK Manifestos

top_words_vec <- quanteda::topfeatures(dfm_mat, 10)

top_words <- tibble::enframe(top_words_vec, name = "feature", value = "frequency")

top_words
```



```{r plot}
## Visualisation

ggplot(top_words, aes(x = reorder(feature, frequency), y = frequency)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Top Words in UK Manifestos",
    x = "Words",
    y = "Frequency"
  )
```
  
## Interpretation

The figure shows the most common words used in UK political manifestos after removing basic words, punctuation, and numbers. Words like “government,” “people,” “labour,” “britain,” and “public” appear most often, which suggests that these texts focus mainly on the state, national identity, and public policy.

The word “labour” appears frequently, which could refer both to the Labour Party and to discussions about jobs and workers. In the same way, words like “government” and “public” show that there is a strong focus on how the country is run and on public services. Words such as “must” and “can” suggest a persuasive style, where parties are making promises and outlining their plans.

Overall, this shows that even a simple word count can help us understand the main topics and priorities in political texts.

Word frequency analysis does help us see the most common topics in the text, but it also has some limitations. Just because a word appears often does not mean we fully understand its meaning or context. Many common words may simply reflect general political language rather than specific policy ideas.

Thus, the analysis could be improved by comparing different political parties (for example, Conservative vs Labour) or looking at how language changes over time. Methods, such as looking at word pairs (bigrams) or identifying themes (topic modelling), could also give a deeper understanding of what the texts are really about.

This pattern reflects how political communication often focuses on broad, shared themes such as governance and public services, rather than highly specific policy language.

## Conclusion

## Conclusion

This analysis demonstrates how text analysis techniques can be used to study large collections of political documents. By transforming text into measurable data, we can identify patterns in political language in a clear and systematic way.

However, simple word frequency analysis has limitations, as it does not capture context or meaning. More advanced methods such as topic modelling or bigram analysis could provide deeper insights into political discourse.
  
  
  






